Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-08-01 paper.tex
Added figures for notations

In our context, split and merge relate to the same dual
problem \todo{not so sure about this actually}. Without loss of
generality, we focus here on merging for the sake of concision. Our
goal then is to merge a set of $n$ chunks into a single reconstructed
3D image with $R$ voxels of size $b$. We assume that all chunks are of
identical size and that slices are squares and blocks are cubes.

@@

We adopt the following notations:
\begin{itemize}
\item $R=D^3$: number of voxels in the reconstructed image.
\item $b$: number of bytes per voxel in the reconstructed image and chunks (in B).
\item $n$: number of chunks (blocks or slices).
\item $m$: amount of available memory (in B).
\item $m'$: amount of used memory (in B), $m'\leq m$.
\end{itemize}

@@

@@

The buffer might be a slightly complex data structure, for instance an
associative array or a Python dictionary, capable of storing multiple
disjoint sequences of contiguous bytes without having to allocate
memory for the bytes between such
sequences. Figure~\ref{fig:cluster-reads-buffer} illustrates how the
buffer would fill up for the two first blocks in a reconstructed
image. Implementation details about the buffer are given later.
\begin{figure}
\centering
\def\svgwidth{0.3\columnwidth}
\input{figures/svg/buffer.pdf_tex}
\caption{Illustration of the buffer used in cluster reads, showing how
  the two first 4x4 blocks of a reconstructed image would be arranged
  in the buffer.  White portions in the buffer are not allocated.  The
  column-major order is used, similar to the NIfTI format: i is the
  fastest changing dimension, followed by j, followed by k.}
\label{fig:cluster-reads-buffer}
% Data file: path
% Gnuplot file: path
% Script file: path
\end{figure}

@@

Finally, the total number of seeks performed by cluster reads to reconstruct the image is:
\begin{equation}
N_\mathrm{CR} =
\begin{cases}
  n + \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{R}^2     & \text{if } \frac{mn}{Rb} < \sqrt[3]{n}\\[0.4em]
  n + \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{R}        & \text{if } \sqrt[3]{n} \leq \frac{mn}{Rb} < \sqrt[3]{n}^2\\[0.4em]
  n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \sqrt[3]{n}^2 \leq \frac{mn}{Rb} < n\\[0.4em]
  n + 1                                                          & \text{otherwise}
\end{cases} \label{eq:seeks-cluster-reads-1}
\end{equation}
It should be noted that $N_{CR}$ is not a continuous function of m,
due to the differences among $b_i$ values.

@@

\todo{Yongping: mention the cache cleaning thing that you run before every experiment.}

@@

(If not added as co-authors): Greg Kiar, Pierre Bellec.

>>>>

In our context, split and merge relate to the same dual problem . We
focus here on merging for the sake of concision. Splitting algorithms
can be derived from merging by swapping reads and writes. Our goal
then is to merge a set of $n$ chunks into a single reconstructed 3D
image with $R$ voxels of size $b$. We assume that all chunks are of
identical size and that slices are squares and blocks are cubes.

@@

We adopt the following notations (see Figure\ref{fig:notations}):
\begin{itemize}
\item $R=D^3$: number of voxels in the reconstructed image.
\item $b$: number of bytes per voxel in the reconstructed image and chunks (in B).
\item $n$: number of chunks (blocks or slices).
\item $m$: amount of available memory (in B).
\item $m'$: amount of used memory (in B), $m'\leq m$.
\end{itemize}

@@

\begin{figure}
  \centering
  \begin{minipage}[b]{0.3\columnwidth}
    \def\svgwidth{\columnwidth}
    \input{figures/svg/Notations.pdf_tex}
    \caption{Notations.}
    \label{fig:notations}
  \end{minipage}
  \quad \quad \quad
  \begin{minipage}[b]{0.3\columnwidth}
    \def\svgwidth{\columnwidth}
    \input{figures/svg/buffer.pdf_tex}
    \caption{Buffer used in cluster reads.}
    \label{fig:cluster-reads-buffer}
  \end{minipage}
\end{figure}

@@

The buffer might be a slightly complex data structure, for instance an
associative array or a Python dictionary, capable of storing multiple
disjoint sequences of contiguous bytes without having to allocate
memory for the bytes between such
sequences. Figure~\ref{fig:cluster-reads-buffer} illustrates how the
buffer would fill up for the two first blocks in a reconstructed
image. It shows how the two first 4x4 blocks of a reconstructed image
would be arranged in the buffer.  White portions in the buffer are not
allocated.  The column-major order is used, similar to the NIfTI
format: i is the fastest changing dimension, followed by j, followed
by k.

@@

Finally, the total number of seeks performed by cluster reads to reconstruct the image is:
\begin{equation}
N_\mathrm{CR} =
\begin{cases}
  n + \ceil*{\frac{Rb}{\sqrt[3]{n}^2m_1'}}\sqrt[3]{R}^2     & \text{if } m < \frac{Rb}{\sqrt[3]{n}^2}\\[0.4em]
  n + \ceil*{\frac{Rb}{\sqrt[3]{n}m_2'}}\sqrt[3]{R}        & \text{if } \frac{Rb}{\sqrt[3]{n}^2} \leq m < \frac{Rb}{\sqrt[3]{n}}\\[0.4em]
  n + \ceil*{\frac{Rb}{m_3'}}                              & \text{if } \frac{Rb}{\sqrt[3]{n}} \leq Rb < n\\[0.4em]
  n + 1                                                          & \text{otherwise}
\end{cases} \label{eq:seeks-cluster-reads-1}
\end{equation}
It should be noted that $N_{CR}$ is not a continuous function of m,
due to the differences among $b_i$ values.

@@

@@

(If not added as co-authors): Greg Kiar, Pierre Bellec, Claude Lepage, Lindsay B. Lewis.
